### 2021-01-12 ch/4-nifr.tex
Finish chapter 5 first draft

\chapter{Null-sampling for Interpretable and Fair Representations}\label{ch:nifr}

>>>>

\chapter{Null-sampling for Interpretable and Fair Representations}\label{ch:nifr}
In the setting from the previous chapter,
labels were untrustworthy because they had been flipped;
a phenomenon we referred to as \emph{label bias}.
However, flipped labels is not the only way that labels can become untrustworthy.
Another way is \emph{sampling bias}, which is the subject of this and the next chapter.

As an example, consider a dataset where the task is to distinguish smiling from non-smiling faces.
Say, we initially have a very diverse and balanced training set,
but then, from the set of smiling faces, we remove nearly all samples where the person does not have red hair,
and from the set of non-smiling faces, we remove nearly all samples with non-black hair.
The result is that even though the labels are still correct, they do not point reliably to the right target anymore.
As hair color is easier to recognize with a \ac{CNN}, the labels have effectively been turned into hair color labels.
In other words, hair color has become a \emph{spurious attribute}.
In the following, we denote the spurious attribute with \(s\),
as it takes a role that is very similar to that of the sensitive attribute that was also denoted by \(s\).

The method, proposed in the previous chapter,
is not able to deal with such a dataset bias as we can easily see:
Say, \emph{smiling} corresponds to \(y=1\) and \emph{not smiling} to \(y=0\);
furthermore, let red hair correspond to \(s=1\), black hair to \(s=0\), and all other hair colors to \(s=2\).
Then, the problem with the described dataset is, that it mostly consists of samples with \(y=0\wedge s=0\)
and those with \(y=1\wedge s=1\).
If we call \(P(y=1|s=s')\) the acceptance rate,
then the problem can be described as one of very different acceptance rates in the hair color groups given by \(s\).
This is the problem tackled in the previous chapter,
and yet, if we were to equalize the acceptance rates with the method there,
the result would be very incorrect.
The issue is that we would treat the labels as incorrect, when in truth, they are correct.

To deal with sampling bias, a different approach is needed.
Indeed, the problem, as posed, is not solvable in the general case.
To make headway with this problem, we introduce the concept of a \emph{context set} (or \emph{pre-training set}).
This set is not subject to the sampling bias,
but is unlabeled (with respect to $y$) and so does not by itself suffice for training.
However, this set does have labels for the spurious attribute \(s\).
This allows us to learn an \emph{invariant representation},
\ie, a representation of the input features \(x\) which is invariant to the spurious attribute.
This kind of representation is equivalent to a fair representation -- as described in \chapref{ch:related-work} --
which is invariant to a sensitive attribute.
With the invariant representation of the training set,
a classifier can then be trained to accurately predict the class label \(y\).

A parallel to the previous chapter is that the method makes use of side information
(in this case the pre-training/context set)
in order to overcome the bias in the training set.

### 2021-01-12 ch/5-zsf.tex
Finish chapter 5 first draft

\chapter{Zero-shot Fairness with Invisible Demographics}\label{ch:zsf}

>>>>

\chapter{Zero-shot Fairness with Invisible Demographics}\label{ch:zsf}
A shortcoming of the approach from the previous chapter
is its reliance on a context/pre-training set which has \(s\) labels.
While it is usually easier to collect data without \(y\) labels and with \(s\) labels,
it is even easier to collect data without any label.
